Static Analysis of Tail Behaviour with a \\ Generalized Gamma Algebra

### 2022-11-23 hta/neurips_2022.tex
Working compile

\documentclass[thesis.tex]{subfiles}
% \documentclass{article}
% \csname remappendix\endcsname

@@

@@

% \begin{abstract}
	Despite the successes of probabilistic models based on passing noise through
	neural networks, recent work has identified that such methods often fail to
	capture tail behavior accurately unless the tails of the base distribution
	are appropriately calibrated. To overcome this deficiency, we propose a
	systematic approach for analyzing the tails of random variables during the
	static analysis (before drawing samples) pass of a probabilistic programming
	language (PPL) compiler. To characterize how the tails change under
	algebraic operations, we develop an algebra acting on a three-parameter
	family of tail asymptotics based on the generalized Gamma distribution. Our
	algebraic operations are closed under addition and multiplication, capable
	of distinguishing sub-Gaussians with differing scales, and handle ratios
	sufficiently well to reproduce the tails of most important statistical
	distributions directly from their definitions. Our experiments confirm that
	inference algorithms leveraging generalized Gamma algebra metadata attain
	superior performance across a number of density modeling and variational
	inference tasks.
% \end{abstract}

@@

\begin{figure}
	\includegraphics[width=\textwidth]{./figures/schematic.pdf}
	\caption{Our overall approach for density approximations with calibrated tails.
		A generative model expressed in a PPL (1) and analyzed using the GGA (2) to
		compute the tail parameters of the target. A representative distribution
		with calibrated tails is chosen for the initial approximation (3) and a learnable
		Lipschitz pushforward (see \Cref{lem:lipschitz}) is optimized (4) to correct the
		bulk approximation.}
\end{figure}

@@

\begin{table}[t]
    \centering
    \bgroup
{\small
\begin{tabular}{|M{0.15\linewidth} | p{0.75\linewidth}|}
\hline
\centering \multirow{2}{*}{\textbf{Ordering}} & 
$\max\{(\nu_1,\sigma_1,\rho_1),(\nu_2,\sigma_2,\rho_2)\}$ \\ & $\qquad \equiv \begin{cases}
(\nu_1,\sigma_1,\rho_1) & \text{ if } \limsup_{x\to\infty} \frac{x^\nu_1 e^{-\sigma_1 x^{\rho_1}}}{x^\nu_2 e^{-\sigma_2 x^{\rho_2}}} < +\infty \\
(\nu_2,\sigma_2,\rho_2) & \text{ otherwise.}
\end{cases}$ \\ \hline
\centering \multirow{2}{*}{\textbf{Addition}} &  $(\nu_{1},\sigma_{1},\rho_{1})\oplus(\nu_{2},\sigma_{2},\rho_{2})$ \\
 & $\equiv \begin{cases}
\max\{(\nu_{1},\sigma_{1},\rho_{1}),(\nu_{2},\sigma_{2},\rho_{2})\} & \text{ if }\rho_{1}\neq\rho_{2}\text{ or }\rho_{1},\rho_{2}<1\\
\left(\nu_{1}+\nu_{2}+1,\min\{\sigma_{1},\sigma_{2}\},1\right) & \text{ if }\rho_{1}=\rho_{2}=1\\
(\nu_{1}+\nu_{2}+\frac{2-\rho}{2},(\sigma_{1}^{-\frac{1}{\rho-1}}+\sigma_{2}^{-\frac{1}{\rho-1}})^{1-\rho},\rho) & \text{ if }\rho=\rho_{1}=\rho_{2}>1.
\end{cases}$ 
\\ \hline
\centering \textbf{Powers} & $(\nu,\sigma,\rho)^\beta \equiv (\frac{\nu+1}{\beta} - 1,\sigma,\frac{\rho}{\beta})$ for $\beta > 0$ \\\hline
\centering \textbf{Reciprocal*} & $(\nu,\sigma,\rho)^{-1} \equiv \begin{cases} (-\nu-2,\sigma,-\rho) & \text{ if } (\nu + 1)/\rho > 0 \text{ and } \rho \neq 0 \\
\mathcal{R}_2 & \text{ otherwise}
\end{cases}$ \\ \hline
\centering \textbf{Scalar \linebreak Multiplication} & $c(\nu,\sigma,\rho) \equiv (\nu,\sigma/|c|^\rho,\rho)$\\\hline
\multirow{3}{*}{\centering \textbf{Multiplication}} & $(\nu_{1},\sigma_{1},\rho_{1})\otimes(\nu_{2},\sigma_{2},\rho_{2})$ \\
&$\qquad \equiv\begin{cases}
\left(\frac{1}{\mu}\left(\frac{\nu_{1}}{|\rho_{1}|}+\frac{\nu_{2}}{|\rho_{2}|}+\frac{1}{2}\right),\sigma,-\frac{1}{\mu}\right) & \text{ if }\rho_{1},\rho_{2}<0\\
\left(\frac{1}{\mu}\left(\frac{\nu_{1}}{\rho_{1}}+\frac{\nu_{2}}{\rho_{2}}-\frac{1}{2}\right),\sigma,\frac{1}{\mu}\right) & \text{ if }\rho_{1},\rho_{2}>0\\
\mathcal{R}_{|\nu_1|} & \mbox{ if }\rho_{1}\leq0,\rho_{2}>0 \\
\mathcal{R}_{\min\{|\nu_1|,|\nu_2|\}} & \mbox{ if }\rho_{1}=0,\rho_{2}=0
\end{cases}$ \\
&where $\mu=\frac{1}{|\rho_{1}|}+\frac{1}{|\rho_{2}|}=\frac{|\rho_{1}|+|\rho_{2}|}{|\rho_{1}\rho_{2}|}$, $\sigma=\mu(\sigma_{1}|\rho_{1}|)^{\frac{1}{\mu|\rho_{1}|}}(\sigma_{2}|\rho_{2}|)^{\frac{1}{\mu|\rho_{2}|}}$.\\ \hline
\centering \textbf{Product of Densities} & $(\nu_1,\sigma_1,\rho_1)\&(\nu_2,\sigma_2,\rho_2) 
\equiv \begin{cases}
(\nu_{1}+\nu_{2},\sigma_{1},\rho_{1}) & \text{ if }\rho_{1}<\rho_{2}\\
(\nu_{1}+\nu_{2},\sigma_{1}+\sigma_{2},\rho) & \text{ if }\rho=\rho_{1}=\rho_{2}\\
(\nu_{1}+\nu_{2},\sigma_{2},\rho_{2}) & \text{ otherwise.}
\end{cases}$ \\ \hline
\centering \textbf{Functions ($L$-Lipschitz)} & 
$f(X_1,\dots,X_n) \equiv L \max\{X_1,\dots,X_n\}$ \\ \hline
    \end{tabular}
}
    \vspace{.25cm}
    \caption{\label{tab:gga_operations}Operations on random variables (e.g.$X_1 + X_2$) are 
    viewed as actions on density functions
    (e.g. convolution $(\nu_1, \sigma_1, \rho_1)\oplus (\nu_2, \sigma_2, \rho_2)$) and the tail parameters of the result are analyzed and reported.}
\egroup
\end{table}

The form of \cref{eq:GenGammaTails} and the name of the algebra is derived from the generalized Gamma distribution. 
\begin{definition}
Let $\nu \in \mathbb{R}$, $\sigma > 0$, and $\rho \in \mathbb{R} \backslash \{0\}$ be such that $(\nu+1)/ \rho > 0$.
A non-negative random variable $X$ is \emph{generalized Gamma distributed} with parameters $\nu,\sigma,\rho$ if it has Lebesgue density
\begin{equation}
\label{eq:GenGammaDensity}
p_{\nu,\sigma,\rho}(x) = c_{\nu,\sigma,\rho} x^\nu e^{-\sigma x^\rho},\qquad x > 0,
\end{equation}
where $c_{\nu,\sigma,\rho} = \rho \sigma^{(\nu+1)/\rho} / \Gamma((\nu+1)/\rho)$ is the normalizing constant. %The generalized Gamma density incorporates many other well-known densities, including that of the Gamma distribution ($\rho = 1$), and the Frechet/Weibull distribution ($\nu = \rho - 1$). 
\end{definition}
The importance of the generalized Gamma form arises due to a combination of two factors:
\begin{enumerate}[label={(\roman*)},leftmargin=*]
    \item The majority of interesting continuous univariate distributions with infinite support satisfy \cref{eq:GenGammaTails}, including
    Gaussians ($\nu=0$, $\rho=2$),
    gamma/exponential/chi-squared ($\nu > -1$, $\rho=1$), Weibull/Frechet ($\rho = \nu + 1$), and
    Student $T$/Cauchy/Pareto ($\mathcal{R}_\nu$).
    However, some notable exceptions include the log-normal distributions. 
    \item The set $\mathcal{G}$ is known to be closed under additive convolution, positive powers, and Lipschitz functions --- we will show it is closed under multiplicative convolution as well. This covers the majority of elementary operations on independent random variables, with reciprocals, exponentials and logarithms the only exceptions. However, we will introduce a few ``tricks'' to handle these cases as well. 
\end{enumerate}
The full list of operations in GGA is compiled in \cref{tab:gga_operations}. All operations in the GGA can be proven to exhibit identical behaviour with their corresponding operations on random variables, with the sole exception of reciprocals (marked by asterisk), where additional assumptions are required. For further details, refer to the Supplementary Material.

	\begin{figure}[h]
		\centering
		\input{./experiments/normal-power-sum.pgf}
		\input{./experiments/lip-nf.pgf}
		\caption{(Top) $5000$ samples of $\lvert x^\top A y \rvert$ vs the
		calibrated GGA density $q(x)$. While calibrated tails are provably guarantees,
		the target distribution's bulk differs from the assumed generalized Gamma representative distribution (\cref{ssec:repr_dist}) for all $k$.
		To fix the bulk approximation, a normalizing flow is composed with
		the GGA representative to form $\text{flow}(q)(x)$. The bulk approximation
		is improved (bottom) while the tails continue to exhibit the same
		behavior (bottom right).
		}
		\label{fig:power_normal}
	\end{figure}

\section{Experiments}\label{sec:experiments}

@@

PPLs can be characterized by the primary use case optimized for, whether that's
Gibbs sampling over Bayes nets \citep{spiegelhalter1996bugs,de2017programming},
stochastic control flow \citep{goodman2012church,wingate2011lightweight},
deep stochastic variational inference \citep{tran2018simple,bingham2019pyro}, or
Hamiltonian Monte-Carlo \citep{carpenter2017stan,xu2020advancedhmc}. Our implementation target
\texttt{beanmachine} \citep{tehrani2020bean} is a declarative PPL selected
due to availability of a PPL compiler and support for static analysis plugins.
Similar to \citet{bingham2019pyro,siddharth2017learning}, it uses PyTorch \citep{paszke2019pytorch} for GPU tensors and automatic differentiation.
Synthesizing an approximating distribution during PPL compilation (\Cref{sec:impl}) is
also performed in the Stan language by \citet{kucukelbir2017automatic} and normalizing
flow extensions in \citet{webb2019aml}. We compare directly against these related density approximators in \Cref{sec:experiments}.

@@

\section{Conclusion}\label{sec:conclusion}

% \item For all authors...
% \begin{enumerate}
%   \item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
%     \answerYes{}
%   \item Did you describe the limitations of your work?
%     \answerYes{See \cref{sec:conclusion}}
%   \item Did you discuss any potential negative societal impacts of your work?
%     \answerYes{See \cref{sec:conclusion}}
%   \item Have you read the ethics review guidelines and ensured that your paper conforms to them?
%     \answerYes{}
% \end{enumerate}

% \item If you ran experiments...
% \begin{enumerate}
%   \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
%     \answerYes{See supplemental, a URL for code will be provided witn camera ready}
%   \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
%     \answerYes{}
% 	\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
%     \answerYes{Standard errors reported in parenthesis on \cref{tab:de} and \cref{tab:vi}}
% 	\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
%     \answerYes{See \cref{sec:experiments}}
% \end{enumerate}

\section{Additional examples}\label{sec:addtl_eg}

@@

\section{Additional details for experiments}

The targets in \Cref{tab:de} and \Cref{tab:vi} are analyzed using
the GGA in \Cref{sec:addtl_eg}. Note that Inverse Gamma (``IG'') corresponds
to the inverse exponential. We selected closed form targets so that
the Pareto tail index $\alpha$ is known analytically and the quality
of theoretical predictions as well as empirical results can be evaluated
against. All experiments are repeated for $100$ trials and $1,000$ samples
from the model (as well as the approximation in VI) were used to compute
each gradient estimate. Losses were trained until convergence, which 
all occurred in under $10,000$ iterations at a $0.05$ learning rate and
the Adam \citep{kingma2014adam} optimizer.

\section{Operations in the Generalized Gamma Algebra}

@@

\paragraph{Addition.} Tails of this form are  closed under addition. Combining subexponentiality for $\rho < 1$
%Any random variable $X$ in $(\nu,\sigma,\rho)$ for $\rho < 1$ is subexponential 
\cite[Chapter X.1]{asmussen2010ruin}, 
%that is, any sum of iid copies of $X$ asymptotically saturates the union bound: $\mathbb{P}\left(\sum_{i=1}^{n}X_{i}>x\right)\sim\sum_{i=1}^{n}\mathbb{P}(X_{i}>x)$ as $x \to \infty$. Combining this
with \cite[Thm 3.1 \& eqn. (8.3)]{asmussen2017tail},
\begin{proposition}
\label{prop:gga_add}
Denoting the addition of random variables (additive convolution of densities) by $\oplus$,
\begin{multline}
(\nu_{1},\sigma_{1},\rho_{1})\oplus(\nu_{2},\sigma_{2},\rho_{2})
\\\equiv \begin{cases}
\max\{(\nu_{1},\sigma_{1},\rho_{1}),(\nu_{2},\sigma_{2},\rho_{2})\} & \text{ if }\rho_{1}\neq\rho_{2}\text{ or }\rho_{1},\rho_{2}<1\\
\left(\nu_{1}+\nu_{2}+1,\min\{\sigma_{1},\sigma_{2}\},1\right) & \text{ if }\rho_{1}=\rho_{2}=1\\
(\nu_{1}+\nu_{2}+1-\frac{\rho}{2},(\sigma_{1}^{-\frac{1}{\rho-1}}+\sigma_{2}^{-\frac{1}{\rho-1}})^{1-\rho},\rho) & \text{ if }\rho=\rho_{1}=\rho_{2}>1.
\end{cases}
\end{multline}
\end{proposition}

@@

\paragraph{Multiplication.}
For any $c \in \mathbb{R} \backslash \{0\}$, it can be readily seen from a change of variables $x \mapsto c x$ that $c (\nu, \sigma, \rho) = (\nu, \sigma / |c|^\rho, \rho)$. The class $\mathcal{G}$ is also closed under multiplication (assuming independence of random variables), as we show in the following result --- the proof is delayed to Appendix C.
\begin{proposition}
\label{prop:Mult}
Denoting the multiplication of independent random variables (multiplicative convolution) by $\otimes$,
\[
(\nu_{1},\sigma_{1},\rho_{1})\otimes(\nu_{2},\sigma_{2},\rho_{2})
\equiv\begin{cases}
\left(\frac{1}{\mu}\left(\frac{\nu_{1}}{|\rho_{1}|}+\frac{\nu_{2}}{|\rho_{2}|}+\frac{1}{2}\right),\sigma,-\frac{1}{\mu}\right) & \text{ if }\rho_{1},\rho_{2}<0\\
\left(\frac{1}{\mu}\left(\frac{\nu_{1}}{\rho_{1}}+\frac{\nu_{2}}{\rho_{2}}-\frac{1}{2}\right),\sigma,\frac{1}{\mu}\right) & \text{ if }\rho_{1},\rho_{2}>0\\
\mathcal{R}_{|\nu_1|} & \mbox{ if }\rho_{1}\leq0,\rho_{2}>0 \\
\mathcal{R}_{\min\{|\nu_1|,|\nu_2|\}} & \mbox{ if }\rho_{1}=0,\rho_{2}=0
\end{cases}
\]
where $\mu=\frac{1}{|\rho_{1}|}+\frac{1}{|\rho_{2}|}=\frac{|\rho_{1}|+|\rho_{2}|}{|\rho_{1}\rho_{2}|}$ and $\sigma=\mu(\sigma_{1}|\rho_{1}|)^{\frac{1}{\mu|\rho_{1}|}}(\sigma_{2}|\rho_{2}|)^{\frac{1}{\mu|\rho_{2}|}}$. 
\end{proposition}
% \begin{proof}
% Appendix C.
% \end{proof}

@@

We can also consider a product of densities operation acting on two random variables $X,Y$, denoted $X \& Y$, by
$p_{X \& Y}(x) = c p_X(x) p_Y(x)$,
where $c > 0$ is an appropriate normalizing constant and $p_X,p_Y,p_{X\& Y}$ are the densities of $X$, $Y$, and $X \& Y$, respectively. In terms of the equivalence classes:
\[
(\nu_1,\sigma_1,\rho_1)\&(\nu_2,\sigma_2,\rho_2) 
\equiv \begin{cases}
(\nu_{1}+\nu_{2},\sigma_{1},\rho_{1}) & \text{ if }\rho_{1}<\rho_{2}\\
(\nu_{1}+\nu_{2},\sigma_{1}+\sigma_{2},\rho) & \text{ if }\rho=\rho_{1}=\rho_{2}\\
(\nu_{1}+\nu_{2},\sigma_{2},\rho_{2}) & \text{ otherwise.}
\end{cases}
\]
Note that this particular operation does not require either $p_X$ or $p_Y$ to be normalized --- only the tail behaviour is needed. We may also use this to work out the tail behaviour of a posterior density, provided the tail behaviour of the likelihood in the parameters is known.

@@

\section{List of univariate distributions}

@@

\caption{List of univariate distributions}\label{tab:dist_list}\\

@@

\section{Proofs}\label{sec:proofs}

>>>>

\documentclass[../thesis.tex]{subfiles}
% \documentclass{article}
% \csname remappendix\endcsname

@@

\chapter{The generalized gamma tail algebra}
\label{ch:hta}

@@

% \begin{abstract}
Whereas previous chapters considered adaptive methods which learn an
approximation's bulk (\cref{ch:lic}) and tail (\cref{ch:ftvi}) from
samples, in this chapter we develop a
systematic approach for analyzing the tails of random variables during the
static analysis (before drawing samples) pass of a probabilistic programming
language (PPL) compiler. To characterize how the tails change under
algebraic operations, we develop an algebra acting on a three-parameter
family of tail asymptotics based on the generalized Gamma distribution. Our
algebraic operations are closed under addition and multiplication, capable
of distinguishing sub-Gaussians with differing scales, and handle ratios
sufficiently well to reproduce the tails of most important statistical
distributions directly from their definitions. Our experiments confirm that
inference algorithms leveraging generalized Gamma algebra metadata attain
superior performance across a number of density modeling and variational
inference tasks. Parts of this chapter have been submitted for peer review
as \fullcite{gga}.
% \end{abstract}

@@

\begin{figure}
  \includegraphics[width=\textwidth]{hta/figures/schematic.pdf}
  \caption{Our overall approach for density approximations with calibrated tails.
    A generative model expressed in a PPL (1) and analyzed using the GGA (2) to
    compute the tail parameters of the target. A representative distribution
    with calibrated tails is chosen for the initial approximation (3) and a learnable
    Lipschitz pushforward (see \Cref{lem:lipschitz}) is optimized (4) to correct the
    bulk approximation.}
\end{figure}

@@

@@

PPLs can be characterized by the primary use case optimized for, whether that's
Gibbs sampling over Bayes nets \citep{spiegelhalter1996bugs,de2017programming},
stochastic control flow \citep{goodman2012church,wingate2011lightweight},
deep stochastic variational inference \citep{tran2018simple,bingham2019pyro}, or
Hamiltonian Monte-Carlo \citep{carpenter2017stan,xu2020advancedhmc}. Our implementation target
\texttt{beanmachine} \citep{tehrani2020bean} is a declarative PPL selected
due to availability of a PPL compiler and support for static analysis plugins.
Similar to \citet{bingham2019pyro,siddharth2017learning}, it uses PyTorch \citep{paszke2019pytorch} for GPU tensors and automatic differentiation.
Synthesizing an approximating distribution during PPL compilation (\Cref{sec:impl}) is
also performed in the Stan language by \citet{kucukelbir2017automatic} and normalizing
flow extensions in \citet{webb2019aml}. We compare directly against these related density approximators in \Cref{sec:hta:experiments}.

@@

\begin{table}[t]
  \centering
    \begin{tabular}{|M{0.15\linewidth} | p{0.75\linewidth}|}
      \centering \multirow{2}{*}{\textbf{Ordering}}        &
      $\max\{(\nu_1,\sigma_1,\rho_1),(\nu_2,\sigma_2,\rho_2)\}$                                                                                                                                                                                                                    \\ & $\qquad \equiv \begin{cases}
          (\nu_1,\sigma_1,\rho_1) & \text{ if } \limsup_{x\to\infty} \frac{x^\nu_1 e^{-\sigma_1 x^{\rho_1}}}{x^\nu_2 e^{-\sigma_2 x^{\rho_2}}} < +\infty \\
          (\nu_2,\sigma_2,\rho_2) & \text{ otherwise.}
        \end{cases}$ \\ \hline
      \centering \multirow{2}{*}{\textbf{Addition}}        & $(\nu_{1},\sigma_{1},\rho_{1})\oplus(\nu_{2},\sigma_{2},\rho_{2})$                                                                                                                                                    \\
                                                           & $\equiv \begin{cases}
                                                                         \max\{(\nu_{1},\sigma_{1},\rho_{1}),(\nu_{2},\sigma_{2},\rho_{2})\}                                              & \text{ if }\rho_{1}\neq\rho_{2}\text{ or }\rho_{1},\rho_{2}<1 \\
                                                                         \left(\nu_{1}+\nu_{2}+1,\min\{\sigma_{1},\sigma_{2}\},1\right)                                                   & \text{ if }\rho_{1}=\rho_{2}=1                                \\
                                                                         (\nu_{1}+\nu_{2}+\frac{2-\rho}{2},(\sigma_{1}^{-\frac{1}{\rho-1}}+\sigma_{2}^{-\frac{1}{\rho-1}})^{1-\rho},\rho) & \text{ if }\rho=\rho_{1}=\rho_{2}>1.
                                                                       \end{cases}$
      \\ \hline
      \centering \textbf{Powers}                           & $(\nu,\sigma,\rho)^\beta \equiv (\frac{\nu+1}{\beta} - 1,\sigma,\frac{\rho}{\beta})$ for $\beta > 0$                                                                                                                  \\\hline
      \centering \textbf{Reciprocal*}                      & $(\nu,\sigma,\rho)^{-1} \equiv \begin{cases} (-\nu-2,\sigma,-\rho) & \text{ if } (\nu + 1)/\rho > 0 \text{ and } \rho \neq 0 \\
              \mathcal{R}_2         & \text{ otherwise}
                                                                                              \end{cases}$                                                                                          \\ \hline
      \centering \textbf{Scalar \linebreak Multiplication} & $c(\nu,\sigma,\rho) \equiv (\nu,\sigma/|c|^\rho,\rho)$                                                                                                                                                                \\\hline
      \multirow{3}{*}{\centering \textbf{Multiplication}}  & $(\nu_{1},\sigma_{1},\rho_{1})\otimes(\nu_{2},\sigma_{2},\rho_{2})$                                                                                                                                                   \\
                                                           & $\qquad \equiv\begin{cases}
                                                                               \left(\frac{1}{\mu}\left(\frac{\nu_{1}}{|\rho_{1}|}+\frac{\nu_{2}}{|\rho_{2}|}+\frac{1}{2}\right),\sigma,-\frac{1}{\mu}\right) & \text{ if }\rho_{1},\rho_{2}<0      \\
                                                                               \left(\frac{1}{\mu}\left(\frac{\nu_{1}}{\rho_{1}}+\frac{\nu_{2}}{\rho_{2}}-\frac{1}{2}\right),\sigma,\frac{1}{\mu}\right)      & \text{ if }\rho_{1},\rho_{2}>0      \\
                                                                               \mathcal{R}_{|\nu_1|}                                                                                                          & \mbox{ if }\rho_{1}\leq0,\rho_{2}>0 \\
                                                                               \mathcal{R}_{\min\{|\nu_1|,|\nu_2|\}}                                                                                          & \mbox{ if }\rho_{1}=0,\rho_{2}=0
                                                                             \end{cases}$                                    \\
                                                           & where $\mu=\frac{1}{|\rho_{1}|}+\frac{1}{|\rho_{2}|}=\frac{|\rho_{1}|+|\rho_{2}|}{|\rho_{1}\rho_{2}|}$, $\sigma=\mu(\sigma_{1}|\rho_{1}|)^{\frac{1}{\mu|\rho_{1}|}}(\sigma_{2}|\rho_{2}|)^{\frac{1}{\mu|\rho_{2}|}}$. \\ \hline
      \centering \textbf{Product of Densities}             & $(\nu_1,\sigma_1,\rho_1)\&(\nu_2,\sigma_2,\rho_2)
      \equiv \begin{cases}
                 (\nu_{1}+\nu_{2},\sigma_{1},\rho_{1})        & \text{ if }\rho_{1}<\rho_{2}      \\
                 (\nu_{1}+\nu_{2},\sigma_{1}+\sigma_{2},\rho) & \text{ if }\rho=\rho_{1}=\rho_{2} \\
                 (\nu_{1}+\nu_{2},\sigma_{2},\rho_{2})        & \text{ otherwise.}
               \end{cases}$                                                                                                                                                                                      \\ \hline
      \centering \textbf{Functions ($L$-Lipschitz)}        &
      $f(X_1,\dots,X_n) \equiv L \max\{X_1,\dots,X_n\}$                                                                                                                                                                                                                            \\ \hline
    \end{tabular}
  \vspace{.25cm}
  \caption{\label{tab:gga_operations}Operations on random variables (e.g.$X_1 + X_2$) are
    viewed as actions on density functions
    (e.g. convolution $(\nu_1, \sigma_1, \rho_1)\oplus (\nu_2, \sigma_2, \rho_2)$) and the tail parameters of the result are analyzed and reported.}
\end{table}

The form of \cref{eq:GenGammaTails} and the name of the algebra is derived from the generalized Gamma distribution.
\begin{definition}
  Let $\nu \in \mathbb{R}$, $\sigma > 0$, and $\rho \in \mathbb{R} \backslash \{0\}$ be such that $(\nu+1)/ \rho > 0$.
  A non-negative random variable $X$ is \emph{generalized Gamma distributed} with parameters $\nu,\sigma,\rho$ if it has Lebesgue density
  \begin{equation}
    \label{eq:GenGammaDensity}
    p_{\nu,\sigma,\rho}(x) = c_{\nu,\sigma,\rho} x^\nu e^{-\sigma x^\rho},\qquad x > 0,
  \end{equation}
  where $c_{\nu,\sigma,\rho} = \rho \sigma^{(\nu+1)/\rho} / \Gamma((\nu+1)/\rho)$ is the normalizing constant. %The generalized Gamma density incorporates many other well-known densities, including that of the Gamma distribution ($\rho = 1$), and the Frechet/Weibull distribution ($\nu = \rho - 1$). 
\end{definition}
The importance of the generalized Gamma form arises due to a combination of two factors:
\begin{enumerate}[label={(\roman*)},leftmargin=*]
  \item The majority of interesting continuous univariate distributions with infinite support satisfy \cref{eq:GenGammaTails}, including
        Gaussians ($\nu=0$, $\rho=2$),
        gamma/exponential/chi-squared ($\nu > -1$, $\rho=1$), Weibull/Frechet ($\rho = \nu + 1$), and
        Student $T$/Cauchy/Pareto ($\mathcal{R}_\nu$).
        However, some notable exceptions include the log-normal distributions.
  \item The set $\mathcal{G}$ is known to be closed under additive convolution, positive powers, and Lipschitz functions --- we will show it is closed under multiplicative convolution as well. This covers the majority of elementary operations on independent random variables, with reciprocals, exponentials and logarithms the only exceptions. However, we will introduce a few ``tricks'' to handle these cases as well.
\end{enumerate}
The full list of operations in GGA is compiled in \cref{tab:gga_operations}. All operations in the GGA can be proven to exhibit identical behaviour with their corresponding operations on random variables, with the sole exception of reciprocals (marked by asterisk), where additional assumptions are required.

\subsection{Illustrative examples}\label{sec:addtl_eg}

@@

\subsection{Operations in the Generalized Gamma Algebra}

@@

\paragraph{Addition.} Tails of this form are  closed under addition. Combining subexponentiality for $\rho < 1$
%Any random variable $X$ in $(\nu,\sigma,\rho)$ for $\rho < 1$ is subexponential 
\cite[Chapter X.1]{asmussen2010ruin},
%that is, any sum of iid copies of $X$ asymptotically saturates the union bound: $\mathbb{P}\left(\sum_{i=1}^{n}X_{i}>x\right)\sim\sum_{i=1}^{n}\mathbb{P}(X_{i}>x)$ as $x \to \infty$. Combining this
with \cite[Thm 3.1 \& eqn. (8.3)]{asmussen2017tail},
\begin{proposition}
  \label{prop:gga_add}
  Denoting the addition of random variables (additive convolution of densities) by $\oplus$,
  \begin{multline}
    (\nu_{1},\sigma_{1},\rho_{1})\oplus(\nu_{2},\sigma_{2},\rho_{2})
    \\\equiv \begin{cases}
      \max\{(\nu_{1},\sigma_{1},\rho_{1}),(\nu_{2},\sigma_{2},\rho_{2})\}                                              & \text{ if }\rho_{1}\neq\rho_{2}\text{ or }\rho_{1},\rho_{2}<1 \\
      \left(\nu_{1}+\nu_{2}+1,\min\{\sigma_{1},\sigma_{2}\},1\right)                                                   & \text{ if }\rho_{1}=\rho_{2}=1                                \\
      (\nu_{1}+\nu_{2}+1-\frac{\rho}{2},(\sigma_{1}^{-\frac{1}{\rho-1}}+\sigma_{2}^{-\frac{1}{\rho-1}})^{1-\rho},\rho) & \text{ if }\rho=\rho_{1}=\rho_{2}>1.
    \end{cases}
  \end{multline}
\end{proposition}

@@

\paragraph{Multiplication.}
For any $c \in \mathbb{R} \backslash \{0\}$, it can be readily seen from a change of variables $x \mapsto c x$ that $c (\nu, \sigma, \rho) = (\nu, \sigma / |c|^\rho, \rho)$. The class $\mathcal{G}$ is also closed under multiplication (assuming independence of random variables), as we show in the following result --- the proof is delayed to Appendix C.
\begin{proposition}
  \label{prop:Mult}
  Denoting the multiplication of independent random variables (multiplicative convolution) by $\otimes$,
  \[
    (\nu_{1},\sigma_{1},\rho_{1})\otimes(\nu_{2},\sigma_{2},\rho_{2})
    \equiv\begin{cases}
      \left(\frac{1}{\mu}\left(\frac{\nu_{1}}{|\rho_{1}|}+\frac{\nu_{2}}{|\rho_{2}|}+\frac{1}{2}\right),\sigma,-\frac{1}{\mu}\right) & \text{ if }\rho_{1},\rho_{2}<0      \\
      \left(\frac{1}{\mu}\left(\frac{\nu_{1}}{\rho_{1}}+\frac{\nu_{2}}{\rho_{2}}-\frac{1}{2}\right),\sigma,\frac{1}{\mu}\right)      & \text{ if }\rho_{1},\rho_{2}>0      \\
      \mathcal{R}_{|\nu_1|}                                                                                                          & \mbox{ if }\rho_{1}\leq0,\rho_{2}>0 \\
      \mathcal{R}_{\min\{|\nu_1|,|\nu_2|\}}                                                                                          & \mbox{ if }\rho_{1}=0,\rho_{2}=0
    \end{cases}
  \]
  where $\mu=\frac{1}{|\rho_{1}|}+\frac{1}{|\rho_{2}|}=\frac{|\rho_{1}|+|\rho_{2}|}{|\rho_{1}\rho_{2}|}$ and $\sigma=\mu(\sigma_{1}|\rho_{1}|)^{\frac{1}{\mu|\rho_{1}|}}(\sigma_{2}|\rho_{2}|)^{\frac{1}{\mu|\rho_{2}|}}$.
\end{proposition}
% \begin{proof}
% Appendix C.
% \end{proof}

@@

We can also consider a product of densities operation acting on two random variables $X,Y$, denoted $X \& Y$, by
$p_{X \& Y}(x) = c p_X(x) p_Y(x)$,
where $c > 0$ is an appropriate normalizing constant and $p_X,p_Y,p_{X\& Y}$ are the densities of $X$, $Y$, and $X \& Y$, respectively. In terms of the equivalence classes:
\[
  (\nu_1,\sigma_1,\rho_1)\&(\nu_2,\sigma_2,\rho_2)
  \equiv \begin{cases}
    (\nu_{1}+\nu_{2},\sigma_{1},\rho_{1})        & \text{ if }\rho_{1}<\rho_{2}      \\
    (\nu_{1}+\nu_{2},\sigma_{1}+\sigma_{2},\rho) & \text{ if }\rho=\rho_{1}=\rho_{2} \\
    (\nu_{1}+\nu_{2},\sigma_{2},\rho_{2})        & \text{ otherwise.}
  \end{cases}
\]
Note that this particular operation does not require either $p_X$ or $p_Y$ to be normalized --- only the tail behaviour is needed. We may also use this to work out the tail behaviour of a posterior density, provided the tail behaviour of the likelihood in the parameters is known.

@@

\subsection{List of univariate distributions}

Here we provide an enumeration of common parametric distributions and their
corresponding GGA parameterizations.

@@

      \caption{List of univariate distributions}\label{tab:dist_list}                                                                                                                                                                                                                                                                         \\

@@

\subsection{Proofs of new results}
\label{sec:proofs}
